% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}


%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
\usepackage{bm}
\usepackage[usenames, dvipsnames]{color}
\usepackage{subcaption}
\captionsetup{compatibility=false}
% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\graphicspath{{./Plots/}}
% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[]{Report Jan 2020 – Attempts at parameterising an empirical AUF}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Tom J. Wilson and Tim Naylor]{
Tom J. Wilson
and Tim Naylor
\\
}

% These dates will be filled out by the publisher
\date{}

% Enter the current year, for the copyright statements etc.
\pubyear{2020}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Given the depths that LSST will reach, we possibly require a different approach than previously undertaken to compute the perturbation component of the Astrometric Uncertainty Function (AUF) for probabilistic photometric catalogue cross-matching. We document here an attempt to derive this AUF -- along with any other non-statistical AUF component, such as proper motions or binary reflex motion -- from the data themselves, creating in-situ AUFs which do not rely on simulation.

\end{abstract}

% to do so we need both cross-matches (using Gaia to have an essentially one-sided problem) to create AUFs for those sources we detect; and could then consider first moments of randomly placed PSF circles where no star exists -- or residuals, if detected source subtracted images were available -- in data frames to get the AUF of stars below the detection limit of the survey

% thus an attempt was made to fit a MoG model to the full cross-match AUF, on the grounds that convolution, and deconvolution, of gaussians is an analytic expression. However, limitations on sigma being >= sigma_statistical (to avoid imaginary numbers in the deconvolution) result in an unphysical perturbation AUF, which may technically work but does not inform about the system at all.

% also explored the attempts to consolidate the number of parameters needed to describe the AUF, by testing whether an effective stellar density could be used, instead of an overall normalising density -- this may work for the very brightest sources where almost all fainter objects in questions are described by a power law (which allows it to work), but very rapidly the turnover in stellar counts renders the tradeoff between central star magnitude and overall local density asymmetric

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The Astrometric Uncertainty Function (AUF) is the probability density function (PDF) describing the probability of a source being detected at some position given its true location. The probability of two sources being the same intrinsic object detected twice given their respective locations -- or the distance between their given positions -- is the convolution of their respective AUFs. Each AUF is, itself, a convolution of the individual components describing its offset from its ``true'' location. The first of these, and typically the only component used, is its statistical component, in which Poisson noise in a photometric image causes the centroiding of a source imperfectly.

A second critical component in crowded regions, or in photometric surveys which probe significant depths relative to the length scale of the survey PSF (of which LSST and \textit{WISE} are both subject), is the perturbation from blended sources. These objects, while hidden beneath the flux from a brighter source, contaminant the central object by both adding additional flux, and by tugging on the center-of-light of the composite object. Thus, in this simple two-component model, the AUF is

\begin{equation}
h(\textbf{x}) = (f_\mathrm{stat} * f_\mathrm{blend})(\textbf{x})
\end{equation}
where the notation $(f * g)(x)$ denotes a convolution.

Previously, \citet{2018MNRAS.481.2148W} produced $f_\mathrm{blend}$ by a Monte Carlo simulation, via Galactic stellar densities from a TRILEGAL simulation \citep{Girardi2005}. However, the limiting magnitude of these simulations -- via the public API -- is 32nd magnitude; this was fine for \textit{WISE} where we still had access to stars 10 magnitudes fainter than the limit of the survey (necessary for probing the faintest flux ratios and sufficiently small flux weighted astrometric perturbations), being brighter than $W1=32$. However, the limiting magnitude of LSST is about 27th magnitude, and thus the magnitude limit is out of the range of the TRILEGAL simulations.

\section{In situ AUF}
Thus the need for deriving empirical AUFs was born. Instead of modelling the perturbations from each component individually -- simulating blended source position shifts, unknown proper motions, etc. -- we could extract from the data themselves the AUF describing \textit{all} components, minus the statistical noise component, and use the data-driven AUF subsequently.

The idea here was that we can reverse the equation above, and compute $f_\mathrm{blend}$ directly from knowing the presciption for the statistical AUF and the total PDF describing the probability of two detections being the same source detected twice given their separation. To this end we matched the \textit{WISE} data, thought to be well understood and accurately described by its quoted uncertainties, to \textit{Gaia} data, again using the high precision on the \textit{Gaia} dataset to effectively reduce the convolution of the two datasets to that of just the \textit{WISE} data, allowing for a one-sided probe of the test.

We are therefore looking for the deconvolution of $f_\mathrm{stat}$ and $f_\mathrm{deblend}$, for which we can turn to Fourier transforms. Where $\mathcal{F}[f(x)](\rho)$ denotes the Fourier transform of $f$, we have

\begin{equation}
	h = f * g
\end{equation}
and, if we take the Fourier transform,

\begin{equation}
H = F \times G
\end{equation}
where in Fourier space the convolution is denoted by a simple multiplication. Thus we can solve for $G$ -- or $f_\mathrm{blend}$ -- as

\begin{equation}
	G = H / F
\end{equation}
and inverse Fourier transform, giving

\begin{equation}
	f_\mathrm{blend}(\textbf{x}) = \mathcal{F}^{-1}\left[\mathcal{F}[h(\textbf{x})](\rho) / \mathcal{F}[f_\mathrm{stat}(\textbf{x})](\rho)\right](\textbf{x}).
\end{equation}

\subsection{Computing the Fourier Transforms}

To reduce the computational expense of a two-dimensional convolution by simplifying it to a 1D integral, we use the Hankel transform,

\begin{equation}
\mathcal{F}[f(r, \theta)](\rho) = \mathcal{F}[f(r)](\rho) = 2 \pi \int\limits_0^\infty\!r f(r) J_0(2 \pi r \rho)\,\mathrm{d}r
\end{equation}
where we assume circular symmetry (i.e., the Hankel transform, or Fourier-Bessel transform, is a two-dimensional Fourier transform under the assumption of a radially symmetric integral kernel) and $J_0$ is the zeroth order Bessel function of first kind. This is computed via a series of small $\Delta r$ bins with each function evaluated at small $r$ (or $\rho$ for an inverse transform) offsets, for a given $\rho$ (or $r$, for inverse transforms).

There is, however, a small instability in this function computed numerically --- no matter how finely you evaluate the functions, $J_0$, effectively a combination of a $1/\sqrt(x)$ function and a $\cos(x)$ function, cannot be reasonably computed with sufficient precision at high $\rho$ in the Fourier transform. At large $\rho$, the $\cos(x)$ component of $J_0$ becomes very rapid, cyling completely every $1/r$ steps (remember it is $J_0(2 \pi r \rho)$ and thus goes as $\cos(2 \pi r \rho)$). While this numerical instability is acceptable for a \textit{convolution} (being of order $10^{-5}$), as a convolution \textit{multiplies} Fourier tranforms, and thus our large scale Fourier transform values, which should be effectively zero, are truncated at $10^{-10}$, and thus do not contribute much when inverse Fourier transformed. Unfortunately, for a \textit{deconvolution}, the functions are \textit{divided}, resulting in an inverse Fourier transform which, for large $\rho$, has a constant \textit{but significant} value -- tending to of order unity. We therefore cannot compute the deconvolution numerically.

\subsection{Mixture of Gaussian Model}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{mnras}
\bibliography{../../../../Dropbox/PostDoc_Exeter/LaTeX/postdoc_papers_jr2}



% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex